---
title: "density statistics"
---
```{r, message=F, echo=F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(emmeans)
library(lme4)

filepath = "C:/Users/madel/OneDrive - NTNU/Doktorgrad/Project 8 Phage therapy and AB in semi-cont/R-analysis/GitHubScripts"
filepath_results = paste0(filepath, "/results/", Sys.Date(), "_")
filepath_figures = paste0(filepath, "/figures/",  Sys.Date(), "_")
source(paste0(filepath, "/plot.settings_waterphage.R"))
flow_density = read.csv(file = paste0(filepath, "/results/2022-11-30_bacterial_density_flow_cyt_clean_nooutliers.csv")) 
bacterial_density = flow_density %>%
  filter(Gate == "Bacteria")
#restructure dataframe
lm_df = bacterial_density %>% 
  dplyr::mutate(Sampling_unit = substring(Sample, 1, 5)) %>%
  dplyr::select(density_cells_mL, DPI, Added_Fc7, Treatment, Sampling_unit) %>%
  dplyr::mutate(Added_Fc7 = factor(Added_Fc7, levels = c("no", "low", "high" )), 
                Treatment = factor(Treatment, levels = c("none", "Phage", "AB")) ) %>%
  as.data.frame()
```

Test for different model fits to the dataset
```{r}
linear_un = lme4::lmer(formula = density_cells_mL ~ DPI * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)
poly2_un = lme4::lmer(formula = density_cells_mL ~ poly(DPI, degree = 2) * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)
poly3_un =lme4::lmer(formula = density_cells_mL ~ poly(DPI, degree =3) * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)
poly4_un = lme4::lmer(formula = density_cells_mL ~ poly(DPI, degree = 4) * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)
linear = lme4::lmer(formula = log10(density_cells_mL) ~ DPI * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)
poly2 = lme4::lmer(formula = log10(density_cells_mL) ~ poly(DPI, degree = 2) * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)
poly3 =lme4::lmer(formula = log10(density_cells_mL) ~ poly(DPI, degree =3) * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)
poly4 = lme4::lmer(formula = log10(density_cells_mL) ~ poly(DPI, degree = 4) * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)
```


```{r}
anova(linear, poly2, poly3, poly4, linear_un, poly2_un, poly3_un, poly4_un)
# the log 10 transformed models are better
AIC = AIC(linear, poly2, poly3, poly4)$AIC

#r sqr val
shap=c(shapiro.test(resid(linear))$p.value, shapiro.test(resid(poly2))$p.value, shapiro.test(resid(poly3))$p.value, 
       shapiro.test(resid(poly4))$p.value)
hist(resid(linear))
hist(resid(poly2))
hist(resid(poly3))
hist(resid(poly4))

plot(poly2, which = 1)
plot(poly3, which = 1)
plot(poly4, which = 1)
qqnorm(resid(poly3))
qqline(resid(poly3))
```

Sticking with the poly3 log10 transformed model

```{r}
library(MuMIn) #install.packages("MuMIn")
r.squaredGLMM(poly3)
```


Predictability is not the best and the residuals are not normally distributed. Thus trying transforming the response variable. 

# Transformations 
To try and get normally distributed residuals I am applying some transformaions on the response variable. 
Both rank and log10 works. 
## log 10 
a transformation that makes sense for density data

```{r}
library(lme4)
linear = lme4::lmer(formula = log10(density_cells_mL) ~ DPI * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)

poly2 = lme4::lmer(formula = log10(density_cells_mL) ~ poly(DPI, degree = 2) * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)

poly3 =lme4::lmer(formula = log10(density_cells_mL) ~ poly(DPI, degree =3) * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)

poly4 = lme4::lmer(formula = log10(density_cells_mL) ~ poly(DPI, degree = 4) * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)


anova(linear, poly2, poly3, poly4)
anova(poly3)
AIC = AIC(linear, poly2, poly3, poly4)$AIC
#r sqr val
shap=c(shapiro.test(resid(linear))$p.value, shapiro.test(resid(poly2))$p.value, shapiro.test(resid(poly3))$p.value, 
       shapiro.test(resid(poly4))$p.value)
hist(resid(linear))
hist(resid(poly2))
hist(resid(poly3))
hist(resid(poly4))

MuMIn::r.squaredGLMM(object = poly3)
```

Will go for log10 since it can be back transformed
Based on investigations of the models I choose to use the log10 transformed 3rd degree polynomial model. 

# poly 3 log10 tranformed model
```{r}
set.seed(3003)
poly3 =lme4::lmer(formula = log10(density_cells_mL) ~ poly(DPI, degree =3) * Treatment * Added_Fc7+ (DPI|Sampling_unit), data = lm_df)
summary(poly3)
anova(poly3)

lm.summary <- summary(poly3)
report = data.frame(Coefficient = rownames(lm.summary$coefficients), 
           Estimate = paste0( format( round(lm.summary$coefficients[,1],1), nsmall = 1 ),  "±", 
                               format( round(lm.summary$coefficients[,2],1), nsmall = 1 ) ), 
           t = format( round(lm.summary$coefficients[,3],1), nsmall = 1 )
           ) 

rownames(report) = NULL

clipr::write_clip(anova(poly3)) #copy to excel table
clipr::write_clip(report) #copy to excel table
```

#Investigate models
```{r}
set.seed(3003)
EMM <-emmeans(object = poly3,  type = "response", adjust = "dunnettx",
        specs = trt.vs.ctrl~ Treatment|poly(DPI, degree = 3)|Added_Fc7, 
        at = list(DPI = c(0,1,2,3,4,5,6,7)))

#get statistics for each comparison per day
EMM$contrasts     # show the means wih dunnettx p-value correction within each day and Fc7 add. 
                  #2 tets per group
EMM$contrasts %>% rbind() #bonferroni correction for all compartisons (42 tests)

estimates_day =

EMM$contrasts %>%#get confidence intervals and p value per day
     summary(infer = TRUE) %>% 
  arrange(contrast) %>% 
  data.frame() %>%
  mutate(across(ratio:t.ratio, round, 3)) %>%
  mutate(p.value = round(p.value,5))

clipr::write_clip(estimates_day) 
```


```{r}
df = emmip(object = poly3, formula =  Treatment|Added_Fc7 ~ poly(DPI, degree = 3),  at = list(DPI = seq(0,7, by = 0.2)), type = "response",
          CIs = T, plotit = F) %>% # which is the same as the inraction term 
  data.frame()%>% 
  rename(density_cells_mL = yvar)

saveRDS(object = df, file = paste0(filepath_results, "lmer_density_estimate_plot.RDS"))
```



