---
title: "ASV table overview"
---
```{r}
library(phyloseq)
library(ggplot2)
library(dplyr)
library(microshades)
```

```{r}
filepath = "C:/Users/madel/OneDrive - NTNU/Doktorgrad/Project 8 Phage therapy and AB in semi-cont/R-analysis/GitHubScripts"
filepath_results = paste0(filepath, "/results/", Sys.Date(), "_")
filepath_figures = paste0(filepath, "/figures/",  Sys.Date(), "_")

source(paste0(filepath, "/plot.settings_waterphage.R"))
source(paste0(filepath, "/ps_melt2.R"))

full_experiment = readRDS(file = paste0(filepath,"/rawdata/Sequencing_Phyloseq-object_16SrDNA.rds") ) 
full_experiment_nocont = subset_samples(full_experiment, Sample_type != "control")
```

#Dataset investigation
```{r}
#full_experiment
#first calculate the maximum sequence depth to test. This is done here by finding the maximum sequence depth #observed in the dataset 
max_depth_to_test = max(phyloseq::sample_sums(full_experiment))

dataset_overview = data.frame(total_reads = sum(phyloseq::sample_sums(full_experiment)), 
           max_read = max(phyloseq::sample_sums(full_experiment)), 
           min_read = min(phyloseq::sample_sums(full_experiment)),
           average_depth = mean(phyloseq::sample_sums(full_experiment)), 
           sd_depth = sd(phyloseq::sample_sums(full_experiment)) )

dataset_overview_nocont = data.frame(total_reads = sum(phyloseq::sample_sums(full_experiment_nocont)), 
           max_read = max(phyloseq::sample_sums(full_experiment_nocont)), 
           min_read = min(phyloseq::sample_sums(full_experiment_nocont)),
           average_depth = mean(phyloseq::sample_sums(full_experiment_nocont)), 
           sd_depth = sd(phyloseq::sample_sums(full_experiment_nocont)) )
```

```{r}
#Plot overview of sampling depth in each sample
df = data.frame(Sample = names(sample_sums(full_experiment)), 
                depth =sample_sums(full_experiment), 
                richness = vegan::renyi(t(as.data.frame(full_experiment@otu_table)), scales = c(0), hill = TRUE),
                sample_data(full_experiment))

ggplot(data = df, aes(x = Sample, y = depth, color = Sample_type)) + 
  geom_point() + 
  my.theme + 
  coord_cartesian(expand = FALSE) + 
  ggh4x::facet_nested(~PP+T, scales = "free_x", space = "free_x") + 
  theme(panel.spacing = unit(0,"lines"))
```
```{r}
df %>% 
  filter(Sample_type == "Water") %>% 
  ggplot(aes(x = depth, y = richness)) + 
  geom_point() 
# no appearent correlation between richness and depth
```

#Composition
Check the composition at genus level of all samples to compare controls and if they make sense. 
```{r, message=FALSE}
# Use microshades function prep_mdf to agglomerate, normalize, and melt the phyloseq object
mdf_rel <- prep_mdf(full_experiment, subgroup_level = "Genus")
abundance_group = mdf_rel %>% group_by(Phylum) %>% 
  summarise(av_ab = mean(Abundance)) %>% 
  arrange(desc(av_ab)) 

#the top 5 phyla contains uncl_d_Bacteria - we dont want this. 
selected_groups_vector = abundance_group$Phylum[c(1:5)]
color_obj_rel = microshades::create_color_dfs(mdf_rel,
                             selected_groups = selected_groups_vector, 
                             group_level = "Phylum",
                             subgroup_level = "Genus", 
                             cvd = TRUE)
mdf_norm_rel <- color_obj_rel$mdf
cdf_norm_rel <- color_obj_rel$cdf
#absolute abundance
mdf_abs <- prep_mdf2(full_experiment, subgroup_level = "Genus")
color_obj_abs = microshades::create_color_dfs(mdf_abs,
                             selected_groups = selected_groups_vector, 
                             group_level = "Phylum",
                             subgroup_level = "Genus", 
                             cvd = TRUE)

mdf_norm_abs <- color_obj_abs$mdf

GP_legend = custom_legend(mdf_norm_rel, legend_key_size = 0.5, legend_text_size =9,
                cdf_norm_rel, 
                group_level = "Phylum",
                subgroup_level = "Genus") + 
  my.theme + 
  theme(panel.border = element_rect(color = "transparent", fill = "transparent"),
        axis.line = element_line(colour = "transparent", linewidth = 0.3))
```

```{r}
rel = microshades::plot_microshades(x = "Sample_name", 
                                         mdf_norm_rel, cdf_norm_rel, group_label = "Phylum Genus") + 
  scale_y_continuous(labels = scales::percent, expand = expansion(0)) +
  my.theme + 
  theme(legend.position = "none",
        text=element_text(size=10), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.title.x = element_blank(), 
        panel.spacing = unit(0.1, "lines")
        ) + 
  ggh4x::facet_nested( ~ Added_Fc7+DPI+Treatment, 
                      scales = "free_x", space = "free_x") + 
  ylab("Relative abundance")

abs = microshades::plot_microshades(x = "Sample_name", 
                                         mdf_norm_abs, cdf_norm_rel, group_label = "Phylum Genus")  +
  my.theme + 
  theme(legend.position = "none",
        text=element_text(size=10), 
        axis.text.x = element_text(angle = 45, size = 5) , 
        axis.ticks.x = element_blank(), 
        axis.title.x = element_blank(), 
        panel.spacing = unit(0.1, "lines")
        ) + 
  ggh4x::facet_nested( ~ Added_Fc7+DPI+Treatment, 
                      scales = "free_x", space = "free_x") + 
  ylab("Sequences")

cowplot::plot_grid( (cowplot::plot_grid(rel, abs, ncol = 1, rel_heights = c(0.45,0.55))),  
                    GP_legend,  rel_widths = c(1, .4))
```

Ordinations with all samples
```{r}
ordination_bray = phyloseq::ordinate(physeq = full_experiment, method = "PCoA", distance = "bray") 
plot_bray = phyloseq::plot_ordination(physeq = full_experiment, ordination = ordination_bray, 
                                      title = "Bray-Curtis",  axes = c(1,2), justDF = T)
evals1 <- round(ordination_bray$values$Eigenvalues[1] / sum(ordination_bray$values$Eigenvalues) * 100, 2)
evals2 <- round(ordination_bray$values$Eigenvalues[2] / sum(ordination_bray$values$Eigenvalues) * 100, 2)

plot_df = plot_bray 
ggplot(plot_df, aes(x=Axis.1, y = Axis.2, fill = Treatment, shape = Sample_type)) + 
  my.theme + 
  ggtitle("Bray-Curtis")  + 
  geom_point(size = 3) + 
  scale_fill_manual(values = c(treatment_cols1, "grey")) + 
  xlab(paste("PCoA 1 (", evals1, "%)", sep = "")) +
  ylab(paste("PCoA 2 (", evals2, "%)", sep = "")) + 
  coord_fixed(ratio = evals2/evals1) + 
  scale_shape_manual(values = c(21,22)) 
```
```{r}
sorensen = vegan::vegdist(t(as.matrix(full_experiment@otu_table@.Data) ), method = "bray", binary = TRUE) #binary as true makes it Sørensen. 
ordination_sor = phyloseq::ordinate(physeq = full_experiment, method = "PCoA", distance = sorensen)
plot_sor = phyloseq::plot_ordination(physeq = full_experiment, ordination = ordination_sor, 
                                   title = "Sørensen", axes = c(1,2), justDF = T) 

evals1 <- round(ordination_sor$values$Eigenvalues[1] / sum(ordination_sor$values$Eigenvalues) * 100, 2)
evals2 <- round(ordination_sor$values$Eigenvalues[2] / sum(ordination_sor$values$Eigenvalues) * 100, 2)

plot_df = plot_sor
ggplot(plot_df, aes(x=Axis.1, y = Axis.2, fill = Treatment, shape = Sample_type)) + 
  my.theme + 
  ggtitle("Sørensen")  + 
  geom_point(size = 3) + 
  scale_fill_manual(values = c(treatment_cols1, "grey")) + 
  xlab(paste("PCoA 1 (", evals1, "%)", sep = "")) +
  ylab(paste("PCoA 2 (", evals2, "%)", sep = "")) + 
  coord_fixed(ratio = evals2/evals1) + 
  scale_shape_manual(values = c(21,22))
```

There is little indication that the sampels are contaminated. The controls are therefore removed for the next steps. 

# Low merging samples
During merging of forward and reverse sequences, some of the samples had below 70% merging. I suspect that some of these samples might not be well sequenced. 

/DNA_17-C-AB-3-D1_S45_L001_R1_001.fastq	63.7% merged
/DNA_66-C-P-2-D3_S17_L001_R1_001.fastq	30.8
/DNA_76-T-AB-2-D3_S126_L001_R1_001.fastq	64.8% merged

```{r}
df.2 = data.frame(Sample = names(sample_sums(full_experiment_nocont)), 
                depth =sample_sums(full_experiment_nocont), 
                richness = vegan::renyi(t(as.data.frame(full_experiment_nocont@otu_table)), scales = c(0), hill = TRUE),
                order1 = vegan::renyi(t(as.data.frame(full_experiment_nocont@otu_table)), scales = c(1), hill = TRUE),
                order2 = vegan::renyi(t(as.data.frame(full_experiment_nocont@otu_table)), scales = c(2), hill = TRUE),
                sample_data(full_experiment_nocont))

df.3 = df.2 %>% 
  mutate(merging = case_when(Sample == "DNA_17.C.AB.3.D1_S45"~ "low", 
                                          Sample == "DNA_66.C.P.2.D3_S17"~ "low", 
                                          Sample == "DNA_76.T.AB.2.D3_S126"~ "low", 
                             TRUE ~ "ok"))

df.3 %>% 
  ggplot(aes(x = T, y  = richness, color = merging)) + 
  geom_point() + 
  ggh4x::facet_nested(~DPI+PP) + 
  my.theme + 
  theme(panel.spacing = unit(0, "lines"))
df.3 %>% 
  ggplot(aes(x = T, y  = order1, color = merging)) + 
  geom_point() + 
  ggh4x::facet_nested(~DPI+PP) + 
  my.theme + 
  theme(panel.spacing = unit(0, "lines"))
df.3 %>% 
  mutate(eveness = order1/richness) %>% 
  ggplot(aes(x = T, y  = eveness, color = merging)) + 
  geom_point() + 
  ggh4x::facet_nested(~DPI+PP) + 
  my.theme + 
  theme(panel.spacing = unit(0, "lines"))
```

# Rarefaction curves
```{r}
sample_data(full_experiment_nocont) = df.3 
#plot rarefraction curves to see differences in sequencing depth
p = ranacapa::ggrare(physeq = full_experiment_nocont, step = 500,  se = FALSE, color = "merging") 
p + 
  my.theme +
  facet_grid(DPI~PP+ T) + 
  ylab("ASV richness") + 
  xlab("Sequencing depth") + 
  theme(text = element_text(size = 14), 
        panel.spacing = unit(0, "lines"))
```

#Removing biased samples
```{r}
full_experiment_nocont = subset_samples(full_experiment, Sample_type != "control") #143 samples
full_experiment_nocont_nobias = subset_samples(full_experiment_nocont, Sequence_name != "DNA_66.C.P.2.D3_S17")
full_experiment_nocont_nobias = subset_samples(full_experiment_nocont_nobias, Sequence_name != "DNA_29.T.N.5.D1_S5")
#141 samples
full_experiment_nocont_nobias = prune_taxa(taxa_sums(full_experiment_nocont_nobias) > 0, full_experiment_nocont_nobias) #4630 taxa
saveRDS(object = full_experiment_nocont_nobias, file = paste0(filepath, "/rawdata/2023-02-02_Phyloseq_object_unnormalised_16srDNA_wo-controls.RDS"))
```



