---
title: "Composition"
---

```{r, message=FALSE}
#libraries
library(dplyr)
library(ggplot2)
library(microshades)
library(cowplot)
```

```{r}
# seed and pathways
set.seed(3003)
filepath = "C:/Users/madel/OneDrive - NTNU/Doktorgrad/Project 8 Phage therapy and AB in semi-cont/R-analysis/GitHubScripts"
filepath_results = paste0(filepath, "/results/", Sys.Date(), "_")
filepath_figures = paste0(filepath, "/figures/",  Sys.Date(), "_")

base::source(paste0(filepath, "/plot.settings_waterphage.R"))
source(paste0(filepath, "/ps_melt2.R"))

relative = readRDS(file = paste0(filepath,"/rawdata/2023-02-02_Phyloseq_object_unnormalised_16srDNA_wo-controls.RDS"))
relative = phyloseq::subset_samples(relative, !Sample_name %in% c("JW-1-D-0", "JW-2-D-0"))
ab_ab = readRDS(paste0(filepath, "/rawdata/2023-02-02_Phyloseq_absolute_abundance_16SrDNA_wo-controls.RDS"))
```

#Order
max 5 selected groups!
The dataframe mdf_norm contains sample data and abundance info. The dataframe cdf_norm stores the color mapping information used for plotting.

plot_microshades produces a stacked barplot with ordered subgroup taxonomy. 
The darkest shade is the most abundant, and the lightest shade is the least abundant. (excluding the “other” subgroup from each of the selected groups)
```{r, message=FALSE}
# Use microshades function prep_mdf to agglomerate, normalize, and melt the phyloseq object
mdf_rel <- prep_mdf(relative, subgroup_level = "Order")
abundance_group = mdf_rel %>% group_by(Phylum) %>% 
  summarise(av_ab = mean(Abundance)) %>% 
  arrange(desc(av_ab)) 

#the top 5 phyla contains uncl_d_Bacteria - we dont want this. 
selected_groups_vector = abundance_group$Phylum[c(1:2,4:6)]
color_obj_rel = microshades::create_color_dfs(mdf_rel,
                             selected_groups = selected_groups_vector, 
                             group_level = "Phylum",
                             subgroup_level = "Order", 
                             cvd = FALSE, 
                             top_orientation = FALSE)

mdf_norm_rel <- color_obj_rel$mdf
cdf_norm_rel <- color_obj_rel$cdf

#Absolute abundance
mdf_abs <- prep_mdf2(ab_ab, subgroup_level = "Order")
color_obj_abs = microshades::create_color_dfs(mdf_abs,
                             selected_groups = selected_groups_vector, 
                             group_level = "Phylum",
                             subgroup_level = "Order", 
                             top_orientation = FALSE,
                             cvd = FALSE)

mdf_norm_abs <- color_obj_abs$mdf
cdf_norm_abs <- color_obj_abs$cdf
mdf_norm_abs = mdf_norm_abs %>% filter(Extraction != "RNA")
#Combinding 
mdf_norm_rel$type = "Relative_abundance"
mdf_norm_abs$type = "Absolute_abundance"

mdf_norm_both = bind_rows(mdf_norm_rel, mdf_norm_abs) %>% 
  dplyr::mutate(type = factor(type, levels = c("Relative_abundance", "Absolute_abundance")), 
                Treatment = case_when(Treatment == "AB"~"A", 
                                      Treatment == "Phage"~"P", 
                                      Treatment == "None"~"N")) %>% 
  dplyr::mutate(Treatment = factor(Treatment, levels = c("N", "P", "A"))) %>%
    dplyr::mutate(Added_Fc7 = case_when(Added_Fc7 == "No" ~ "No-Fc7", 
                                      Added_Fc7 == "Low" ~ "Low-Fc7", 
                                      Added_Fc7 == "High" ~ "High-Fc7")) %>%
  mutate(Added_Fc7 = factor(Added_Fc7, levels = c("No-Fc7", "Low-Fc7", "High-Fc7" )))
```


```{r, message=FALSE}
GP_legend = custom_legend(mdf_norm_rel, legend_key_size = 0.5, legend_text_size =25,
                cdf_norm_rel, 
                group_level = "Phylum",
                subgroup_level = "Order") + 
  my.theme + 
  theme(panel.border = element_rect(color = "transparent", fill = "transparent"),
        axis.line = element_line(colour = "transparent", linewidth = 0.3))

#Absolute and relative abundance
rel_abs_plot =

microshades::plot_microshades(x = "Sample_name", 
                              mdf_norm_both, cdf_norm_rel, 
                              group_label = "Phylum Order") + 
  my.theme + 
  coord_cartesian(expand = FALSE) + 
  theme(legend.position = "none",
        text=element_text(size=25), 
        axis.text.x = element_blank() , 
        axis.ticks.x = element_blank(), 
        axis.title.x = element_blank(), 
        strip.text= element_text(size = 22),
        panel.spacing = unit(0, "lines"), 
        panel.grid.major.x = element_blank(),
        axis.title.y = element_blank()
        ) + 
  ggh4x::facet_nested(type ~ Added_Fc7+DPI+Treatment, 
                      scales = "free", space = "free_x", switch = "both") 

cowplot::plot_grid(rel_abs_plot, GP_legend,  rel_widths = c(1, .4))
```


```{r}
ggsave(filename = paste0(filepath_figures, "Figure5_Relative-Absolute-abundance_Phylum-Order.pdf"), 
       width = 180*3.5, height = 120*2.5, units = "mm", dpi = 600)
```




