---
title: "Differential absolute abundance analysis"
---

```{r, message=FALSE}
library(ggplot2)
library(phyloseq)
library(dplyr)
library(tidyr)
library(corncob) #devtools::install_github("bryandmartin/corncob", dependencies = TRUE, force = TRUE)
```

# Filter dataset

```{r}
filepath = "C:/Users/madel/OneDrive - NTNU/Doktorgrad/Project 8 Phage therapy and AB in semi-cont/R-analysis/GitHubScripts"
filepath_results = paste0(filepath, "/results/", Sys.Date(), "_")
filepath_figures = paste0(filepath, "/figures/",  Sys.Date(), "_")
source(paste0(filepath, "/plot.settings_waterphage.R"))
```

```{r}
#select taxa to use for DA analysis. Criteria is that the total abundanance of the ASV is over 1% (0.01 in rel abundance) and has a prevalance in over 5% of the samples (i.e. 8 samples)
ps_exp_comp = readRDS(file = paste0(filepath,"/rawdata/2023-02-02_Phyloseq_absolute_abundance_16SrDNA_wo-controls.RDS"))
ps_exp_comp_r_filt = metagMisc::phyloseq_filter_prevalence(ps_exp_comp, 
                                      prev.trh = 0.05, 
                                      abund.trh = 2500, 
                                      abund.type = "total",
                                      threshold_condition = "AND")


#1508 taxa used for DA analysis
```

```{r}
taxa_sums_ps = taxa_sums(ps_exp_comp) %>% data.frame() %>% 
         rename(Total = ".") %>% arrange(desc(Total)) %>%
  filter(Total< 5000) %>%
  mutate(cutoff = case_when(Total < 2500 ~ "remove", T ~"keep"))
  
ggplot(data =taxa_sums_ps, aes(x=Total, fill = cutoff)) + 
  geom_histogram(bins = 200) + 
  scale_y_continuous(trans = "log10")
```

```{r}
#start with raw dataset. algoritms are fixing the sampling bias 
ps_exp = ps_exp_comp_r_filt
d7 = phyloseq::subset_samples(physeq = ps_exp, DPI == "7")
sample_data(d7) = sample_data(d7) %>% data.frame() %>% 
  mutate(group = paste0(sample_data(d7)$Treatment, "_", sample_data(d7)$Added_Fc7)) %>% 
  mutate(group = factor(group, levels = c("None_No", "None_Low", "None_High",
                                          "Phage_No","Phage_Low", "Phage_High", 
                                          "AB_No", "AB_Low", "AB_High"))) 
d7_DNA = prune_taxa(taxa_sums(d7) > 0, d7) #1505 taxa 7DPI, 1508 ASVs total
saveRDS(d7_DNA, file = paste0(filepath_results, "Phyloseq_absolute_abundance_16SrDNA_7DPI-filtered-1508ASVs.RDS"))
```

We are interested in testing for differential taxa wihin each propagule pressure, comparing phage and AB to the control treatment

#Corncob


The corncob package (Douglas,et al https://doi.org/10.1038/s41467-022-28034-z</div>)

We converted the metadata and non-rarefied feature tables into a phy-loseq object (version 1.29.0)75, which we input to corncob’s differentialTest func-tion (version 0.1.0)10. This functionfits each taxon count abundance to a beta-binomial model, using logit link functions for both the mean and overdispersion.

Because corncob models each of these simultaneously and performs both differ-ential abundance and differential variability testing10, we set the null over-dispersion model to be the same as the non-null model so that only taxa havingdifferential abundances were identified. 

Finally, the function performed sig-nificance testing, for which we chose Wald tests (with the default non-bootstrapsetting), and we obtained BH FDR-correctedp-values as output

```{r}
d7_DNA_no = subset_samples(d7_DNA, Added_Fc7 == "No") 
da_analysis_DNA_n = differentialTest(formula =  ~ Treatment, #only interested in differential abundance  
                               phi.formula = ~  Treatment, 
                               formula_null = ~ 1, 
                               phi.formula_null = ~ Treatment,
                               test = "Wald", #wald significance test 
                               boot = FALSE, #default to false 
                               data = d7_DNA_no,#the phyloseq object 
                               fdr_cutoff = 0.05) #alpha value for significance
## low DNA
d7_DNA_low = subset_samples(d7_DNA, Added_Fc7 == "Low") 
da_analysis_DNA_l = differentialTest(formula =  ~ Treatment, #only interested in differential abundance  
                               phi.formula = ~  Treatment, 
                               formula_null = ~ 1, 
                               phi.formula_null = ~ Treatment,
                               test = "Wald", #wald significance test 
                               boot = FALSE, #default to false 
                               data = d7_DNA_low,#the phyloseq object 
                               fdr_cutoff = 0.05) #alpha value for significance
#high
d7_DNA_high = subset_samples(d7_DNA, Added_Fc7 == "High") 
da_analysis_DNA_h = differentialTest(formula =  ~ Treatment, #only interested in differential abundance  
                               phi.formula = ~  Treatment, 
                               formula_null = ~ 1, 
                               phi.formula_null = ~ Treatment,
                               
                               test = "Wald", #wald significance test 
                               boot = FALSE, #default to false 
                               data = d7_DNA_high,#the phyloseq object 
                               fdr_cutoff = 0.05) #alpha value for significance
```

```{r}
df = plot(da_analysis_DNA_n, data_only = T, level = "ASV")  %>% mutate(Added_Fc7 ="No")
corncob_result_df = df
corncob_list = list(da_analysis_DNA_n = da_analysis_DNA_n)

df = plot(da_analysis_DNA_l, data_only = T,level = "ASV") %>% mutate(Added_Fc7 ="Low")
corncob_result_df = bind_rows(df, corncob_result_df)
list = list(da_analysis_DNA_l = da_analysis_DNA_l)
corncob_list = append(corncob_list, list)

df = plot(da_analysis_DNA_h, data_only = T, level = "ASV") %>% mutate(Added_Fc7 ="High")
corncob_result_df = bind_rows(df, corncob_result_df)
list = list(da_analysis_DNA_h = da_analysis_DNA_h)
corncob_list = append(corncob_list, list)
#refine dataframe
plotdf = corncob_result_df %>% 
  separate(variable, sep = "\n",
           into =  c("Treatment", "Dif_abd"  ) ) %>% 
  mutate(Treatment = substring(Treatment, 10)) %>% 
  mutate(ASV = taxa) %>% 
  dplyr::rename(b_estimate = x, CI_low95 =xmin, CI_high95 = xmax )
ASV_tax = data.frame(tax_table(d7)) %>% select(Domain:ASV) %>% distinct() %>% filter(ASV %in% plotdf$ASV)
plotdf2 = merge(x = plotdf,y =ASV_tax, by = "ASV")

corncob_list = append(corncob_list, list(df_results = plotdf2))
saveRDS(object = corncob_list, file = paste0(filepath_results, "Corncob_list-results_ASV_DNA_absolute_1508.RDS"))
```

# ANCOMBC

prv_cut	- a numerical fraction between 0 and 1. Taxa with prevalences less than prv_cut will be excluded in the analysis. For instance, suppose there are 100 samples, if a taxon has nonzero counts presented in less than 10 samples, it will not be further analyzed. Default is 0.10.

lib_cut	-a numerical threshold for filtering samples based on library sizes. Samples with library sizes less than lib_cut will be excluded in the analysis. Default is 0, i.e. do not discard any sample.


```{r}
# ANCOM
set.seed(3003)
d7_fam <- d7_DNA
d7_fam@sam_data$tr_group = d7_fam@sam_data$group
#ancombc also supports importing data in phyloseq format
out_tr_add = ANCOMBC::ancombc2(#tax_level = NULL, aggregates at lowest tax rank - i.e. ASV
              data = d7_fam, #phyloseq object
              fix_formula  = "tr_group", 
              p_adj_method = "holm", 
              prv_cut = 0, #Taxa with prevalences less than prv_cut will be excluded in the analysis 
              lib_cut = 0, #no samples are removed
              group = "tr_group",
              struc_zero = TRUE, 
              pseudo = 0, 
              pseudo_sens = TRUE,
              pairwise = T, #compare all groups to each other
              neg_lb = F, 
              iter_control = list(tol = 0.01, max_iter = 100, verbose = FALSE),
              alpha = 0.05,
              n_cl = 1, verbose = TRUE)

saveRDS(object = out_tr_add, file = paste0(filepath_results, "ANCOMBC_results_Group-treatments_ASV_DNA_absolute_1508.RDS"))

##  Structural zeros
output = out_tr_add
tab_zero = output$zero_ind
res_pair = output$res_pair
res_pair %>%
    dplyr::select(taxon, contains("diff")) 
# we want to compare pairwise within the added Fc7 groups. no added control vs phage, no added control vs ab, low added ....
df_fig =
res_pair %>%
  #keep taxa that are differentially abundant in groups of interest
    dplyr::filter(diff_tr_groupPhage_No == 1 | 
                    diff_tr_groupAB_No == 1| 
                    diff_tr_groupPhage_Low_tr_groupNone_Low == 1| 
                    diff_tr_groupAB_Low_tr_groupNone_Low == 1| 
                    diff_tr_groupPhage_High_tr_groupNone_High == 1| 
                    diff_tr_groupAB_High_tr_groupNone_High == 1) %>%
  #change value to 0 if not significantly different in the group
  dplyr::mutate(lfc_tr_groupPhage_No = ifelse(diff_tr_groupPhage_No == 1, 
                                              lfc_tr_groupPhage_No, 0),
                lfc_tr_groupAB_No = ifelse(diff_tr_groupAB_No == 1, 
                                           lfc_tr_groupAB_No, 0),
                lfc_tr_groupPhage_Low_tr_groupNone_Low = 
                  ifelse(diff_tr_groupPhage_Low_tr_groupNone_Low == 1, 
                         lfc_tr_groupPhage_Low_tr_groupNone_Low, 0),
                lfc_tr_groupAB_Low_tr_groupNone_Low = 
                  ifelse(diff_tr_groupAB_Low_tr_groupNone_Low == 1, 
                         lfc_tr_groupAB_Low_tr_groupNone_Low, 0),
                lfc_tr_groupPhage_High_tr_groupNone_High = 
                  ifelse(diff_tr_groupPhage_High_tr_groupNone_High == 1, 
                         lfc_tr_groupPhage_High_tr_groupNone_High, 0),
                lfc_tr_groupAB_High_tr_groupNone_High = 
                  ifelse(diff_tr_groupAB_High_tr_groupNone_High == 1, 
                         lfc_tr_groupAB_High_tr_groupNone_High, 0)) %>% 
  #change names to more readable names
  dplyr::transmute(taxon, 
              `No_added Phage vs control` = round(lfc_tr_groupPhage_No, 2),
              `No_added AB vs control` = round(lfc_tr_groupAB_No, 2), 
              `Low_added Phage vs control` = round(lfc_tr_groupPhage_Low_tr_groupNone_Low, 2),
              `Low_added AB vs control` = round(lfc_tr_groupAB_Low_tr_groupNone_Low, 2),
              `High_added Phage vs control` = round(lfc_tr_groupPhage_High_tr_groupNone_High, 2),
              `High_added AB vs control` = round(lfc_tr_groupAB_High_tr_groupNone_High, 2))   %>% 
  #restructure df to have groups in one column and values in another instead of a matrix format
    pivot_longer(cols = `No_added Phage vs control`:`High_added AB vs control`, 
                 names_to = "group", values_to = "value") %>%
    arrange(taxon) %>% #arrange by alphabetical order
    mutate(comparison = factor(group, levels = c("No_added Phage vs control", 
                                                 "Low_added Phage vs control",   
                                                 "High_added Phage vs control" ,
                                                 "No_added AB vs control", 
                                                 "Low_added AB vs control" ,
                                                 "High_added AB vs control" )))

gen_df = data.frame(tax_table(d7_DNA)) %>% select(Domain:ASV) %>% distinct() %>% 
  filter(ASV %in% df_fig$taxon) %>% mutate(taxon = ASV)
df_fig_t = merge(x = df_fig,y = gen_df, by = "taxon")
saveRDS(object = df_fig_t, file = paste0(filepath_results, "ANCOMBC_result_pairwize-organised_ASV_DNA_absolute_1508.RDS"))
```

# DeSeq2
Following guide given in Statistical Analysis of Microbiome Data with R chapter 11.5 The DESeq and DESeq2 Packages https://doi.org/10.1007/978-981-13-1534-3

```{r}
library(DESeq2) # BiocManager::install("DESeq2")
library(phyloseq)
library(dplyr)
library(ggplot2)
library(tidyr)
# DeSeq2
set.seed(3003)
physeq = d7_DNA
# convert OTU table to matrix
otu_tab = (otu_table(physeq))
#DESeq2 need taxa(genes=rows) by samples(=columns)
countData<-(as(otu_tab, "matrix"))
#head(countData)
#Step 2: Create the Sample Metadata Table
meta = data.frame(sample_data(physeq))
group = meta$group
#Step 3: Build the DESeq2 Object
# build the deseq object 
metaData = data.frame(row.names=colnames(countData),group=group)
dds = DESeqDataSetFromMatrix(countData = countData, colData = metaData, design = ~group)
dds <- dds[rowSums(counts(dds)) > 0,] #double check all 0 count taxa are gone
#normalise data with median ration method.
#Step 5: Normalize the Count Data
dds <- estimateSizeFactors(dds)
#sizeFactors(dds)
#Step 6: Estimate the Dispersion
#estimate the dispersion 
dds<- estimateDispersions(dds)
#Step 7: Test the Differential Abundance
deseq_restult = DESeq2::DESeq(object = dds, 
                              test = "Wald", 
                              fitType = "parametric" )
saveRDS(object = deseq_restult, file = paste0(filepath_results, "DeSeq_object_ASV_DNA_absolute_1508.RDS"))

res <- results(deseq_restult, contrast = c("group", "Phage_No", "None_No"))
df = as.data.frame(res) 
df$taxa = rownames(df) 
rownames(df) = NULL
df = na.omit(df) #only keep significant taxa
df = df %>% mutate(Added_Fc7 = "No", 
              Treatment = "Phage", 
              Extraction = "DNA", 
              method = "DeSeq2") %>% 
  filter(padj < 0.05)
df1 = df

res <- results(deseq_restult, contrast = c("group", "Phage_Low", "None_Low"))
df = as.data.frame(res) 
df$taxa = rownames(df) 
rownames(df) = NULL
df = na.omit(df) #only keep significant taxa
df = df %>% mutate(Added_Fc7 = "Low", 
              Treatment = "Phage", 
              Extraction = "DNA", 
              method = "DeSeq2") %>% 
    filter(padj < 0.05)
df1 = bind_rows(df1, df)

res <- results(deseq_restult, contrast = c("group", "Phage_High", "None_High"))
df = as.data.frame(res) 
df$taxa = rownames(df) 
rownames(df) = NULL
df = na.omit(df) #only keep significant taxa
df = df %>% mutate(Added_Fc7 = "High", 
              Treatment = "Phage", 
              Extraction = "DNA", 
              method = "DeSeq2") %>% 
    filter(padj < 0.05)
df1 = bind_rows(df1, df)

res <- results(deseq_restult, contrast = c("group", "AB_No", "None_No"))
df = as.data.frame(res) 
df$taxa = rownames(df) 
rownames(df) = NULL
df = na.omit(df) #only keep significant taxa
df = df %>% mutate(Added_Fc7 = "No", 
              Treatment = "AB", 
              Extraction = "DNA", 
              method = "DeSeq2") %>% 
    filter(padj < 0.05)
df1 = bind_rows(df1, df)

res <- results(deseq_restult, contrast = c("group", "AB_Low", "None_Low"))
df = as.data.frame(res) 
df$taxa = rownames(df) 
rownames(df) = NULL
df = na.omit(df) #only keep significant taxa
df = df %>% mutate(Added_Fc7 = "Low", 
              Treatment = "AB", 
              Extraction = "DNA", 
              method = "DeSeq2") %>% 
    filter(padj < 0.05)
df1 = bind_rows(df1, df)

res <- results(deseq_restult, contrast = c("group", "AB_High", "None_High"))
df = as.data.frame(res) 
df$taxa = rownames(df) 
rownames(df) = NULL
df = na.omit(df) #only keep significant taxa
df = df %>% mutate(Added_Fc7 = "High", 
              Treatment = "AB", 
              Extraction = "DNA", 
              method = "DeSeq2") %>% 
    filter(padj < 0.05)
df1 = bind_rows(df1, df)

#Create a dataframe with the significant OTUs
alpha = 0.05
sigtab = df1[which(df1$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(physeq)[sigtab$taxa, ], "matrix"))
# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))
sigtab2 = sigtab %>% mutate(classified = substr(Order, 1,1)) %>% 
  mutate(Ordername =case_when(classified == "o" ~ substring(Order, 3), 
                              classified != "o" ~ "x_unclassified at order level")) %>% 
  mutate(classified2 = substr(Genus, 1,1)) %>% 
  mutate(taxonname = case_when(classified2 == "g" ~ substring(Genus,1), 
                              classified2 != "g" ~  substring(Genus, 6) )) %>%
  mutate(lowest_class = substr(taxonname, 1,1)) %>%
  mutate(group = 
           case_when(classified2 == "g" ~  paste0(substring(taxonname,3)), 
                     classified2 != "g" ~  paste0("uncl. (", substring(taxonname,1,1), " ", substring(taxonname,3), ")"))) %>%
  mutate(classified_c = substr(Class, 1,1)) %>% 
  mutate(classname =case_when(classified_c == "c" ~ substring(Class, 3), 
                              classified_c != "c" ~ "x_unclassified at class level")) %>%
  mutate(classified_f = substr(Family, 1,1)) %>% 
  mutate(classname_f =case_when(classified_f == "f" ~ substring(Family, 3), 
                              classified_f != "f" ~ "x_unclassified at family level"))

write.csv(x = sigtab2, file = paste0(filepath_results, "DeSeq_result-dataframe_ASV_DNA_absolute_1508.csv"))
```